---
title: "Biliary Drainage and Stenting: A service evaluation"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc-depth: 3
    toc-float: true
    theme: flatly
    embed-resources: true
knitr:
  opts_chunk:
    echo: false
    warning: false
    message: false
    fig-width: 9
    fig-height: 5
    out-width: "100%"
---

```{r setup}
#| include: false
source("sources/libraries.R")
source("sources/dataWrangle.R")
source("sources/descriptiveStats.R")
source("sources/mortality.R")
source("sources/survival.R")
source("sources/regression.R")
source("sources/machineLearning.R")
```

## Background

[Brief background on biliary drainage and stenting — indications, clinical context, institutional setting.]

## Methods

Data were collected retrospectively for all biliary drainage and stenting procedures performed at Glasgow Royal Infirmary (GRI) and the Queen Elizabeth University Hospital (QEUH) between `r study_start` and `r study_end`. [Include ethics/audit registration as appropriate.]

Continuous variables are presented as median (IQR) and compared using the Mann-Whitney U test. Categorical variables are presented as n (%) and compared using the chi-squared test. Survival analysis was performed using the Kaplan-Meier method with log-rank testing; multivariable survival analysis used Cox proportional hazards regression. Machine learning models (logistic regression, random forest, XGBoost) were trained on the earliest 80% of procedures and validated on the most recent 20% (temporal validation).

## Results

### Patient cohort

A total of `r nrow(biliary)` procedures were performed across the study period — `r sum(biliary$Site == "GRI")` at GRI and `r sum(biliary$Site == "QEUH")` at QEUH, averaging `r avg_per_month_total` procedures per month overall. `r sum(biliary$Diagnosis == "Malignant")` procedures (`r round(mean(biliary$Diagnosis == "Malignant") * 100)`%) were for malignant obstruction and `r sum(biliary$Diagnosis == "Benign")` (`r round(mean(biliary$Diagnosis == "Benign") * 100)`%) for benign obstruction.

#### Procedures per year

```{r chart-procs-total}
chart_procs_total
```

```{r chart-procs-site}
chart_procs_site
```

#### Patient demographics

```{r chart-age}
#| fig-height: 4
chart_age
```

### Baseline characteristics

#### By hospital

```{r table1a}
table1a
```

#### By sex

```{r table1b}
table1b
```

#### By diagnosis

```{r table1c}
table1c
```

### Mortality

#### 30-, 42-, and 90-day mortality

```{r table-mortality}
table_mortality
```

```{r chart-mortality}
chart_mortality
```

#### Inpatient mortality

```{r table-inpatient}
table_inpatient
```

#### Mortality by age decade (malignant obstruction)

```{r table-decade-mortality}
table_decade_mortality
```

### Survival analysis

Median overall survival was `r median_overall` days.

#### Overall survival

```{r surv-overall}
#| fig-height: 7
surv_overall
```

#### Survival by diagnosis

```{r surv-diagnosis}
#| fig-height: 7
surv_diagnosis
```

The following analyses are restricted to the malignant obstruction cohort (n = `r nrow(biliary_malignant)`). The x-axis is limited to `r median_overall` days (median overall survival) to focus on the clinically relevant early period.

#### Survival by sex

```{r surv-sex}
#| fig-height: 7
surv_sex
```

#### Survival by hospital

```{r surv-hospital}
#| fig-height: 7
surv_hospital
```

#### Survival by prior ERCP

```{r surv-ercp}
#| fig-height: 7
surv_ercp
```

### Cox proportional hazards regression

Multivariable Cox regression was performed in the malignant obstruction cohort (n = `r nrow(biliary_cox_complete)`). Weight was excluded due to incomplete data (~`r round(mean(is.na(biliary_malignant$Weight)) * 100)`% missing).

```{r tbl-cox-merged}
tbl_cox_merged
```

```{r chart-cox-forest}
chart_cox_forest
```

### Machine learning: 42-day mortality prediction

Three models were trained to predict 42-day mortality using temporal validation (earliest 80% training, most recent 20% test).

```{r table-ml-binary}
table_ml_binary
```

#### ROC curves

```{r chart-roc}
chart_roc_binary
```

#### Variable importance (random forest)

```{r chart-vip}
#| fig-height: 6
chart_vip_rf
```

#### Model calibration

```{r chart-calibration}
chart_calibration
```

## Discussion

[Discussion of key findings.]

## Conclusion

[Conclusions.]
